\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

Large language models (LLMs) have been shown to be persuasive in debate and conversational
formats, but there is growing evidence that LLMs can also act as persuasive debaters or
conversational partners capable of moving people's stances on complex, value-laden topics
in relatively brief interactions. Moreover, an understudied question concerns whether humans
can similarly cause AI systems to adapt their argumentation across
the course of a debate. This paper proposes a debate-format study using GPT-4o mini,
measuring within-subject attitude and confidence shifts in human participants and recorded belief
changes in AI agents across debate rounds, and probing the impact of assumed human versus
artificial intelligence (AI) identity modulating trust and openness to influence.
The primary empirical analysis focuses on AI-to-human persuasion: how distinct AI personalities
and rhetorical strategies influence human belief and confidence outcomes.
A key asymmetry is that human participants complete between-round reflections paraphrasing their
opponent's argument, whereas AI agents do not; reflection quality is therefore analyzed exclusively
as a human mediator of persuasion.

\section*{Key Findings}

Across $n = 89$ participants in four debate conditions (human-human control, firm AI, balanced AI, open-minded AI),
we find a striking dissociation between belief and confidence outcomes:

\begin{itemize}
    \item \textbf{Null Belief Change}: AI debate produces no significant effect on participant belief relative to
          human-human debate across four pre-specified confirmatory models (all $p > 0.29$, $d = 0.31$).
    \item \textbf{Robust Post-Debate Confidence Difference}: Despite null belief change, AI debate significantly increases
          post-debate participant confidence ($\beta = 12$--$13$ points, $95\%$ CI: $[3, 24]$, Bonferroni $p = 0.02$--$0.024$).
          Baseline confidence was recorded as 0 for all participants, so this reflects post-debate level differences.
    \item \textbf{Mechanism: Illusory Success}: Participants who engaged critically with the AI (high pushback)
          showed \textit{lower} confidence gains ($r = -0.73$, $p = 0.007$), contradicting a ``defensive confidence'' hypothesis.
          Confidence gains reflect perceived debate success rather than substantive persuasion.
\end{itemize}

\section*{Contributions}

This thesis contributes to understanding AI persuasiveness by demonstrating that:
(1) the confidence-belief dissociation is reproducible and theoretically novel;
(2) AI debate is more effective for participant self-efficacy and argumentation practice than for genuine belief change;
(3) rigorous small-sample inference with locked designs and multiplicity correction is feasible in human-AI interaction research.

This study is well aligned with and extends this emerging literature on AI-mediated persuasion and
human-AI interaction dynamics, offering implications for AI's role in education, argumentation training, and collaborative decision-making.