\section{Conclusion}

Building on Chapter 5's synthesis, this thesis evaluates AI persuasiveness in a debate platform,
with confirmatory comparison of pooled AI opponents against human-human debate.
The primary empirical finding is that confirmatory belief and confidence effects were not identified
at this sample size because confidence intervals include zero. This pattern is
methodologically informative,
offering implications for AI's role in persuasion, argumentation training, and human-AI collaboration.

\subsection{Primary Finding: Directional Confidence-Belief Dissociation Pattern}

Across four pre-specified confirmatory models with robust standard errors,
AI debate yielded belief estimates with confirmatory 95\% CIs that include 0
($p > 0.35$, $d = 0.286$, observed power $= 20.0\%$).
Confidence models showed a consistent directional increase with 95\% CIs that also include 0
($\beta = 5.2$--$5.8$ points, $p = 0.17$--$0.21$).

Baseline confidence was not degenerate in this run (0.0\% at floor and 20.0\% at ceiling at the Round 1 baseline),
so confidence analyses remain interpretable as change outcomes though the 20\% ceiling at the Round 1 baseline may compress
upward variance and bias estimates toward the null — the directional increase observed ($\beta = 5.2$--$5.8$) should therefore be
treated as a lower-bound signal rather than a precise effect estimate.

This directional pattern is consistent with classical persuasion theory, which often models belief and
confidence as closely coupled constructs \cite{petty1986_elm,eagly1993_psychology_of_attitudes}.
In the AI debate context, participants may update their \textit{confidence in debate performance}
without updating their \textit{core stance conviction}. They finish debates feeling more certain
in direction, without reliable evidence of shifted underlying belief. Because both confirmatory
outcomes have CIs that include zero, this is interpreted as a directional dissociation \textit{pattern}
rather than a confirmed dissociation effect.

\subsection{Exploratory Mechanism Hypothesis: Pushback and Confidence}

I examined whether the directional confidence pattern reflected a "defensive confidence" mechanism:
participants who argued critically should feel more convinced of their position after "winning" debate exchanges.
Correlation analysis of the qualitative subsample suggested a possible negative association:
participants who engaged in high pushback (critical argumentation) showed a numeric trend toward lower confidence gain
($r = -0.298$, $p = 0.229$, 95\% CI $[-0.672,0.196]$). Jackknife resampling
was used to check sensitivity to influential observations and yielded a directional negative range
($r_{jackknife} \in [-0.412,-0.169]$), with substantial small-sample spread.
One speculative interpretation is \textit{illusory success}: some participants may have interpreted
surface-level debate fluency as personal success, increasing confidence without corresponding belief revision.
Given $n=18$ in the mechanism subset, this mechanism result is hypothesis-generating,
small-sample, and not effect-identifying.

\subsection{Theoretical Implications}

\subsubsection{Possible Interpretation: AI as Intellectual Interlocutor Without Epistemic Authority}

Participants may treat AI as a serious intellectual opponent (as suggested by continued engagement and post-debate attribution responses),
yet remain reluctant to grant AI epistemic authority sufficient to shift core beliefs. This tentative interpretation suggests:

\begin{enumerate}
      \item Opponent-attribution responses indicate that many AI-condition participants reported suspecting AI identity,
              which is consistent with at least partial source-based discounting on
          value-laden or identity-central topics.
    \item Confidence gains are divorced from belief shifts, perhaps because confidence is self-referential
          ("How sure am I of my position?") while belief resistance reflects source skepticism ("How credible is the AI?").
    \item For topics where AI might have genuine epistemic authority (factual matters, logical consistency),
          the debate platform may not capture belief shifts because participant stance choices are often taken on priors.
\end{enumerate}

\subsubsection{Implications for AI Pedagogy and Argumentation Training}

If AI debate directionally boosts confidence without shifting beliefs, its value may lie in \textit{self-efficacy training}
rather than \textit{persuasion}. Participants may benefit from:
\begin{itemize}
    \item Increased confidence in their own reasoning and argumentation skills,
    \item Reduced anxiety about defending positions,
    \item Practice formulating counterarguments in real time.
\end{itemize}

However, if the intent is to expose participants to genuinely compelling opposing views and shift deeply held beliefs,
the current pilot evidence does not support a reliable belief-shift effect from AI debate alone.

\subsection{Methodological Strengths}

This analysis documents an explicit small-sample inference workflow through:
\begin{itemize}
      \item \textbf{Pre-specified model sequence}: Four confirmatory models specified before final model execution.
    \item \textbf{Robust standard errors}: HC3 covariance matrices account for heteroscedasticity.
      \item \textbf{Inferential transparency}: confidence models are reported with explicit effect sizes,
                    confidence intervals, and corresponding $p$-values.
                                    \item \textbf{Transparency about underpowering}: With model sample $n = 89$, observed power for belief effect detection is $20.0\%$.
                                                                              The CI-over-zero pattern is stable across reported sensitivity checks.
      \item \textbf{Mechanism sensitivity audit attempted}: exploratory correlation and leave-one-out checks
                    are reported transparently as small-sample diagnostics rather than confirmatory mechanism evidence.
      \item \textbf{Data quality audits}: Floor-ceiling effect checks and baseline confidence distribution diagnostics.
\end{itemize}

\pagebreak
\section{Limitations}

\subsection{Sample Size and Power}

With model sample $n = 89$, this study is underpowered for \textit{belief} effects ($d = 0.286$).
With the actual complete-case allocation ratio of approximately 1:3.45 (human-human $n = 20$, pooled AI $n = 69$), achieving 80\% power at this effect size
would require approximately $n \approx 124$ for human-human and $n \approx 429$ for pooled AI (total $n \approx 553$).
The unequal group allocation — a consequence of non-randomised assignment — means the human-human cell ($n = 20$)
is the binding constraint on power, not the pooled AI cell ($n = 69$).
The CI-over-zero pattern is stable across robustness checks, and replication with a larger sample is required.
The descriptive cohort is $n = 90$; one case is omitted from regression models under complete-case
rules due to variable-level missingness, yielding analytic $n = 89$.

\subsection{Non-Randomized Assignment and Residual Confounding}

Condition assignment was administratively balanced rather than fully randomized. This improves
practical cell balance in a pilot, but weakens strict causal identification because assignment may
correlate with unobserved factors (for example, participation timing, topic-selection patterns,
or latent debate motivation). Topic fixed effects and baseline covariates reduce but do not eliminate
this vulnerability. Therefore, condition contrasts in this pilot should be read as associational with
controlled adjustment, not as definitive randomized-treatment effects.

\subsection{Measurement and Scale Issues}

\begin{itemize}
    \item \textbf{Belief scale coarseness}: 0--100 continuous scale may not capture nuanced within-debate shifts,
          particularly for topics with high pre-engagement ceiling effects.
    \item \textbf{Confidence endpoint uncertainty}: Confidence baselines are usable in this run,
          and confirmatory confidence intervals still include 0.
    \item \textbf{Secondary-model residual non-normality}: Residual normality checks failed for
          the confidence-model family (conf-M1--conf-M3), so confidence-family estimates
          are treated as cautious, distribution-sensitive summaries in this pilot.
    \item \textbf{Reflection quality measurement limits}: The main reflection predictor is a rubric-based GPT composite.
          Text length was retained only as a fallback diagnostic and was not used in the final confirmatory models;
          ideally, qualitative codes would be applied exhaustively rather than to a stratified sample.
\end{itemize}

\subsection{Robustness and Model Sensitivity}

Planned robustness checks include outlier trimming, leave-one-out conditional re-estimation, VIF diagnostics, and residual normality assessment.
\begin{itemize}
    \item One pre-specified robustness check (leave-one-condition) produced no output in the current analysis run and is therefore omitted from inference.
\end{itemize}

\subsection{Sample Representation Bias}

Qualitative analysis was performed on a stratified sample ($n \approx 24$ reflections)
that in the current export is closer to population condition shares
(human-human $25.0\%$ vs. $20.5\%$; firm $25.0\%$ vs. $28.4\%$; balanced $25.0\%$ vs. $27.4\%$;
open-minded $25.0\%$ vs. $23.7\%$).
Representativeness risk is reduced relative to earlier skewed splits, but mechanism findings remain exploratory,
hypothesis-generating, small-sample, and not effect-identifying.

\subsection{Generalizability}

Participants were recruited based on availability and volunteered into debate participation, possibly skewing toward
debate-curious, relatively confident individuals. Results may not generalize to the broader population.
Additionally, topics were participant-selected rather than experimenter-controlled, introducing unmeasured heterogeneity.

\subsection{AI Realism and Authenticity}

The AI opponent is a prompted language model, not a persistent agent with learning. The pattern of
prompt-constrained response behavior, while methodologically clean, may not reflect authentic disagreement dynamics
in real-world human-human or human-AI debate.

\pagebreak
\section{Future Work}

Future research directions address the dissociation finding and explore boundary conditions:

\subsection{Belief Attribution Separately Randomized}

Implement a $2 \times 2$ design crossing actual opponent type (AI vs. human) with \textit{belief about} opponent type.
This directly addresses the central source-attribution confound and isolates whether confidence effects arise
from real AI properties or from participants' \textit{belief} that they debated an AI.

\subsection{Longitudinal Belief Tracking}

Does the confidence boost persist, and does it eventually translate into belief shifts?
Measuring participant beliefs days or weeks post-debate, with follow-up booster debates,
could reveal whether initial illusory confidence resolves into genuine persuasion with repeated exposure.

\subsection{Factual vs. Value-Based Topics}

The current sample spans diverse topics. Stratified analysis by topic type
could test whether AI is more persuasive on factual matters (where it has genuine epistemic advantage)
than value-laden stances (where source credibility matters more). The current
platform topic set used in this thesis is listed in Appendix A,
Section~\ref{sec:appendix_topic_questions}.

\subsection{Qualitative Deepening and Full Coding}

Apply thematic coding exhaustively to all reflections and arguments, not just a stratified sample.
Integrate quantitative outcome measures with qualitative content analysis to triangulate mechanism hypotheses.

\subsection{AI Personality Fine-Tuning}

Current AI personalities (firm, balanced, open-minded) show no effect on belief outcomes.
However, given the pilot sample size and the study's broader underpowered status,
I cannot distinguish between genuine absence of personality effects and insufficient power to detect them.
Experiment with AI personalities that explicitly concede strategic points, express uncertainty, or
adopt Socratic questioning. Would such modifications increase persuasiveness?
Current baseline prompt templates are documented in Appendix A,
Section~\ref{sec:appendix_ai_personality_prompts}.

\subsection{AI as Knowledge Authority vs. Opinion Debater}

Separate debates into two tracks: (1) factual knowledge debates (e.g., scientific claim verification)
where AI has genuine advantage, and (2) value/opinion debates where human credibility dominates.
Hypothesize that AI persuasiveness is moderated by topic type.

\subsection{Practical Application: Argumentation Tutoring}

Since AI debate shows directional confidence increases, deploy the platform as a confidence-building tool in
argumentation curricula. Compare students' debate performance pre- and post-AI-debate practice.
Test whether confidence gains transfer to performance against human opponents.

\pagebreak
\section{Final Remarks}

This thesis contributes to our understanding of AI persuasiveness by demonstrating that
AI in debate format does not produce reliable pooled effects on either belief change or confidence change
in this pilot, while still supporting useful engagement and process diagnostics.
This pattern motivates a nuanced model of AI's role:
not as a replacement for human-led persuasion or education, but as a complement for practice,
confidence building, and intellectual sparring.

The methodological framework---locked designs, robust inference, and mechanism validation---
provides a template for rigorous empirical work on AI-human interaction in persuasive contexts.
As AI becomes increasingly integrated into learning and decision-support systems,
understanding its limitations and comparative advantages remains critically important.
Numeric claims in this chapter are synchronized to Appendix
Table~\ref{tab:headline_numbers}.
