\section{Overview}

This chapter describes the experimental methodology for examining AI persuasiveness
in a turn-based debate format. The primary analysis focuses on \textbf{AI-to-human
persuasion}: whether debating against a GPT-4o mini opponent configured with one of
three personalities (firm, balanced, open-minded) produces different belief outcomes
relative to a human-human control condition. A secondary analysis
investigates confidence change as a secondary outcome. Exploratory analyses examine
reflection quality as a mechanism and argument quality as an alternative predictor.
This study is a \textbf{pilot investigation}; effect sizes are estimated and
infrastructure for future full-scale deployment is established. Confirmatory claims
are not made at this sample size.

A key structural asymmetry exists in the design: human participants complete
between-round reflections (paraphrase and acknowledgement), whereas AI agents do not.
Reflection quality is therefore analyzable only on the human side. Analysis of
human-to-AI persuasion is reserved for future work.


\section{Study Design}

The study uses a controlled between-subjects design with repeated measures over
debate rounds. Participants are assigned to one of four conditions: three AI opponent
conditions (firm, balanced, open-minded) or a human-human control. Each debate
follows an identical round structure with arguments submitted in alternating turns
and belief and confidence checks collected after each round, with confidence first collected at
the first post-argument round (Round 1).


\section{Participants and Recruitment}

Participants were recruited according to institutional guidelines and provided with
an informed consent statement describing the debate task and data usage. Inclusion
required completion of the full debate protocol, including per-round belief and
confidence checks from Round 1 onward and the post-debate survey. The final analytic sample comprised
$n = 90$ participants across four conditions: human-human ($n = 21$), firm AI
($n = 28$), balanced AI ($n = 20$), and open-minded AI ($n = 21$).


\section{Conditions and Assignment}

Participants were assigned to conditions with an operational goal of maintaining
approximate balance across the three AI personalities and the human-human control.
Assignment was monitored through an administrative matchmaking workflow and adjusted
during data collection to keep condition counts as balanced as feasible. Opponent
identity was concealed during the debate to prevent source bias. Topic selection
and initial stance were self-selected by participants; some imbalance across
conditions on these dimensions may remain and is addressed by including topic
fixed effects in the primary models.
Condition assignment was not formally randomized; participants were allocated to
maintain approximate balance via administrative monitoring.


\section{Procedure}

\begin{enumerate}
      \item Participant selects a debate topic and indicates initial stance.
    \item Participant completes pre-debate survey recording openness to persuasion:
          \textit{firm on my stance}, \textit{convinced of my stance}, or
          \textit{open to changing my mind}.
    \item Debate proceeds in alternating argument rounds (500-character limit per
          turn). After each round, the participant records updated belief value
          (0--100); confidence (0--100) is recorded from the first post-argument
          round onward.
    \item Between rounds, the participant completes a reflection task comprising
          a free-text paraphrase of the opponent's most recent argument and a
          free-text acknowledgement of whether they concede the opponent's point.
    \item After the final round, the participant completes a post-debate survey
          including opponent attribution and confidence in that attribution.
\end{enumerate}

\pagebreak
\section{Measures}

\subsection{Measurement Strategy Rationale}

The measurement design intentionally combines round-level process tracking with
debate-level endpoint self-reports. Persuasion in live debate is expected to be
path-dependent: participants may temporarily shift, rebound, or stabilize as new
arguments are introduced. Collecting belief, confidence, and perceived opponent
influence after each round captures these within-debate trajectories rather than
only net change. The reflection and acknowledgement prompts add a complementary
mechanism layer by recording whether participants can paraphrase opponent content
and explicitly concede valid points, which helps distinguish superficial numeric
movement from substantive engagement.

Pre-debate and post-debate self-reports serve a different function. The pre-debate
item anchors baseline openness to persuasion, while post-debate items capture the
participant's global interpretation of debate impact and attribution judgments.
Together, this process-plus-endpoint strategy provides convergent evidence across
temporal, cognitive, and self-evaluative dimensions, improving interpretability
relative to single-timepoint designs.

\subsection{Primary Outcome: Belief Change}

Participant belief is captured on a continuous 0--100 scale, recorded before the
debate (Round 0) and after each subsequent round. Belief change is computed as
$\Delta \text{belief} = \text{belief}_{\text{post}} - \text{belief}_{\text{pre}}$,
where pre-debate belief is taken from Round 0 and post-debate belief from the
participant's final round response.

\subsection{Secondary Outcome: Confidence Change}

Confidence is measured on a 0--100 scale at the first post-argument round
(Round 1; confidence was not collected at pre-debate Round 0, so Round 1 is used as \texttt{confidence\_pre},
noting that this is a post-exposure baseline --- the participant has already seen the opponent's first argument ---)
and at the participant's final round (\texttt{confidence\_post}).
Confidence change is computed as
$\Delta \text{confidence} = \text{confidence}_{\text{post}} - \text{confidence}_{\text{pre}}$.
Baseline confidence at Round 1 shows substantial variation (M = 79.3, SD = 18.4),
but with ceiling effects \(20\%\) at \(100\), so confidence-change models are
interpreted with this constrained variance in mind.
Specifically, a 20\% ceiling at 100 at the Round 1 baseline may compress upward variance and
bias confidence-change estimates toward the null.

\subsection{Floor/Ceiling Audit}

Floor/ceiling checks were examined during measurement setup to interpret scale
constraints before model fitting. In the current sample, belief pre-debate values
show floor/ceiling mass (20.0\% at 0; 16.7\% at 100), while confidence at the Round 1 baseline
shows 0.0\% at floor and 20.0\% at ceiling. These distributions support using change-score
models while requiring caution for confidence-change inference under upper-end compression.
For endpoint context, post-debate values show lower boundary concentration
(belief post: 11.1\% at 0, 3.3\% at 100; confidence post: 0.0\% at 0, 14.4\% at 100).
The observed confidence-to-belief variance ratio is 0.297, consistent with lower
confidence-change variance and reduced power to detect confidence effects.

\subsection{Reflection Quality}

Between-round reflections were scored on an overall quality scale
using GPT-4o mini as the primary rater, drawing on dimensions of
faithfulness, specificity, engagement, clarity, and charity
collapsed into a single composite score. To validate GPT scoring,
two independent human raters scored a stratified random subsample
of $n = 30$ reflections using the same rubric. Convergent validity
was assessed using Spearman $\rho$ between the GPT score and the
mean of the two human ratings; inter-rater reliability between
human raters was assessed using two-way random effects ICC$(2,1)$
(single-rater reliability) and ICC$(2,k)$ (average-rater
reliability), implemented via mean-squares decomposition. A
combined ICC was additionally computed by treating GPT as a third
rater alongside the two humans. Rows with missing values on any
rater were excluded prior to ICC computation. Reflection quality
scores were standardized ($z$-scored) prior to inclusion as a
model covariate.

\subsection{Argument Quality}

Persuader argument quality was scored as a single overall quality
rating using GPT-4o mini, drawing on dimensions of evidence
quality, logical coherence, responsiveness, nuance, and clarity.
Validation followed an identical procedure to reflection quality:
two independent human raters scored a stratified subsample of
$n = 30$ arguments; Spearman $\rho$ was computed between GPT
scores and the mean human rating; two-way random effects
ICC$(2,1)$ and ICC$(2,k)$ were computed for human-only and
GPT$+$human matrices respectively, with listwise deletion of
incomplete rows. Argument quality scores were standardized
($z$-scored) prior to inclusion as an exploratory covariate.

\subsection{Pre-Debate Openness}

Participants completed a single-item pre-debate survey selecting one of three
openness options. This variable is treated as a baseline covariate
(\texttt{initial\_stance\_strength\_z}) capturing resistance to persuasion and
is included in confirmatory models M4 and M5.

\subsection{Acknowledgement}

The free-text acknowledgement field records whether the participant explicitly
conceded validity in the opponent's argument. This was binarized
(\texttt{acknowledgement\_binary}) and included as an exploratory covariate in
a secondary belief change model (\texttt{belief\_ack}).
The binary coding rule marked acknowledgement as 1 when responses included explicit
concession terms such as "agree", "valid point", "fair point", "you are right",
"I concede", or close paraphrases; otherwise it was coded 0. Coding was applied by
the primary researcher using this fixed keyword-and-phrase rule and then checked
against random examples during data cleaning.

\subsection{Opponent Attribution}

The post-debate survey records opponent perception (human, AI, or unsure),
perception confidence (1--5), suspicion timing (round range), detection cues
(multi-select), and AI awareness effect (more persuasive, less persuasive, no
difference). These variables are available for exploratory analysis but are
not part of the primary confirmatory model sequence.

\subsection{Survey Instrument Wording (Pre and Post)}

To make the measurement instrument explicit, this subsection reproduces the
participant-facing pre-debate and post-debate survey items used in the platform.

\paragraph{Pre-debate survey item}

Prompt shown before matching:
\begin{quote}
\textit{Tell us how you currently feel about your stance on this topic.}
\end{quote}

Response options:
\begin{itemize}
      \item \textit{I am firm on my stance.}
      \item \textit{I am convinced of my stance.}
      \item \textit{I am open to changing my mind about this.}
\end{itemize}

\paragraph{Post-debate survey items}

After debate completion, participants answer the following items:
\begin{enumerate}
      \item \textit{Overall, how did this debate affect your original stance on this topic?}
      \\
      Options: \textit{Strengthened my original view}; \textit{Slightly strengthened};
      \textit{No real effect}; \textit{Slightly weakened}; \textit{Weakened my original view}.

      \item \textit{How strong is your stance on this topic now?}
      \\
      7-point scale from \textit{Very weak (1)} to \textit{Very strong (7)}.

      \item \textit{How confident are you in your stance now?}
      \\
      7-point scale from \textit{Not confident at all (1)} to
      \textit{Extremely confident (7)}.

      \item \textit{Do you think your opponent was AI or human?}
      \\
      Options: \textit{Human}; \textit{AI}; \textit{Unsure}.

      \item \textit{How confident are you in your answer above?}
      \\
      5-point scale from \textit{Not confident at all (1)} to
      \textit{Extremely confident (5)}.

      \item \textit{When did you first suspect your opponent was [AI/human/unclear]?}
      \\
      Options: \textit{Round 1--2}; \textit{Round 3--4}; \textit{Round 5--7};
      \textit{Round 8--12}; \textit{Round 13--17}; \textit{Round 18--20};
      \textit{I never suspected / realized only now}.

      \item \textit{What made you think your opponent was [AI/human/unclear]?}
      \\
      Multi-select cue checklist with an \textit{Other (please specify)} field.

      \item Conditional awareness item (shown based on response to Item 4):
      \begin{itemize}
            \item If \textit{AI}: \textit{Did knowing your opponent was AI affect your stance change?}
            \item If \textit{Human}: \textit{Did knowing your opponent was human affect your stance change?}
            \item If \textit{Unsure}: \textit{Did your uncertainty about your opponent's nature affect your stance change?}
      \end{itemize}
      Options: \textit{No difference}; \textit{Less persuasive};
      \textit{More persuasive}; \textit{Not sure / can't say}, plus a required
      short free-text justification.
\end{enumerate}


\section{Data Quality and Exclusion Criteria}

Debates were excluded if incomplete, missing per-round belief checks, or
containing inconsistent round sequencing. The descriptive sample is $n = 90$,
while complete-case model analyses use $n = 89$ due to one case with
variable-level missingness in model inputs. No implausible
response time exclusions were applied; all exclusions are reported
transparently.

\pagebreak
\section{Statistical Analysis}

This study was not pre-registered; as a substitute discipline, the model sequence,
confirmatory/exploratory split, and interpretation hierarchy were specified before final model execution.

\subsection{Pre-Model Checks}

Prior to confirmatory model fitting, four diagnostic checks were
conducted.
First, zero-mass on \texttt{belief\_change} was audited
overall (14.4\%) and by condition (human-human: 23.8\%; firm: 10.7\%;
balanced: 15.0\%; open-minded: 9.5\%).
Second, a Shapiro-Wilk test
confirmed approximate normality of \texttt{belief\_change}
($W = 0.980$, $p = 0.190$), supporting OLS use.
Third, a one-way ANOVA indicated no significant between-condition
difference in \texttt{belief\_change} prior to covariate adjustment ($F = 0.488$,
$p = 0.691$); an AI-subcondition Kruskal-Wallis check yielded $\eta^2 = 0.001$
($H = 0.606$, $p = 0.739$), supporting pooling as a pilot-stage precision choice
rather than proving personality equivalence.
Fourth, topic fixed effects were justified by an adjusted
$R^2$ increment of 0.070 over condition-only models, confirming topics
are a meaningful confounder. An \texttt{nrounds} imbalance was noted:
the balanced condition averaged 5.80 rounds vs.\ 4.19 for human-human
($SD = 2.44$ vs.\ $0.51$), motivating its inclusion as a covariate in
M5.

\subsection{Rationale for Pooling AI Conditions}

The three AI personality conditions (firm, balanced, open-minded) are pooled
into a single binary indicator (\texttt{group\_2lvl}: AI vs.\ human-human)
for the primary confirmatory analysis. This decision is motivated by two
defensible considerations. First, the pilot sample ($n = 89$) yields per-condition
cell sizes of $n = 20$--$28$, which are insufficient to reliably estimate
separate personality effects; pooling improves precision for the
higher-priority human-vs-AI contrast, while remaining underpowered for
small effects in this pilot. Second, the primary research question
concerns whether AI debate \textit{per se} produces different outcomes than
human debate, independent of personality configuration; personality-level
differences are a secondary question addressed in Section~\ref{sec:part10}.
A null ANOVA result under low power does not confirm personality equivalence.
Condition-level trajectories are examined in the
four-level analysis (Section~\ref{sec:part10}), which retains the full
condition structure.


\subsection{Primary Model Sequence: Confirmatory and Embedded Exploratory Models}

Before model estimation, the confirmatory hypotheses were:
\begin{itemize}
    \item \textbf{H1 (belief outcome):} Participants in pooled AI conditions differ from
          participants in the human-human condition on belief change.
    \item \textbf{H2 (confidence outcome):} Participants in pooled AI conditions differ from
          participants in the human-human condition on confidence change.
\end{itemize}

Belief change is modelled using OLS regression with explicit covariance settings.
For models without topic fixed effects (M1--M3), OLS-HC3
standard errors are used. For models with topic fixed effects (M4, M4c, M5),
topic-cluster standard errors are used to align inference with the
topic-level grouping structure. The focal predictor is a binary group indicator
(AI pooled vs.\ human-human), with human-human as the reference category.
Seven models are estimated (four confirmatory, three exploratory/sensitivity):

OLS-HC3 is used as the primary estimator because the confirmatory estimands
are mean-difference contrasts under likely heteroskedasticity in a pilot sample.
As a methods-justification sensitivity
check (not a primary finding), separate OLS-HC3/RLM-Huber/two-stage WLS
refits are compared for the same primary model family in Appendix
Table~\ref{tab:estimator_comparison_primary}.

\begin{itemize}
    \item \textbf{M1}: Treatment effect only — unadjusted estimate of AI
          vs.\ human-human belief change.
    \item \textbf{M2}: Treatment $+$ reflection quality ($z$-scored) —
          adjusting for participant engagement depth.
     \item \textbf{M3} (exploratory): Treatment $\times$ reflection quality
          interaction — testing whether reflection quality moderates the
          effect of AI debate on belief change.
    \item \textbf{M4}: Treatment $+$ reflection quality $+$ initial stance
          strength ($z$-scored) $+$ topic fixed effects — the primary
          fully-adjusted confirmatory model.
    \item \textbf{M4b} (exploratory): Treatment $+$ reflection quality $+$
          persuader argument quality ($z$-scored) — an omitted-variable
          check testing whether argument quality accounts for variance
          beyond reflection quality alone, without topic or stance controls.
    \item \textbf{M4c} (exploratory): Treatment $+$ persuader argument
          quality $+$ initial stance strength $+$ topic fixed effects —
          substituting argument quality for reflection quality to isolate
          the independent predictive contribution of persuader argument
          quality.
    \item \textbf{M5}: Full model — treatment $+$ reflection quality $+$
          argument quality $+$ initial stance strength $+$ topic fixed
          effects $+$ number of rounds $+$ first-player indicator, serving
          as a kitchen-sink sensitivity check for coefficient stability.
\end{itemize}

Given the pilot model sample ($n = 89$),
observed power for the M1 treatment effect was 0.20 at the observed
Cohen's $d = 0.286$. With the actual complete-case allocation ratio of approximately 1:3.45
(human-human $n = 20$, pooled AI $n = 69$),
power is constrained by the smaller cell; to achieve 80\% power at this effect size and allocation ratio
would require approximately $n \approx 124$ for human-human and $n \approx 429$ for pooled AI (total $n \approx 553$).
The observed power estimate of 20\% is computed at the full model $n = 89$,
treating the two-group contrast as the unit of analysis; the unequal group sizes
(AI n = 69, human-human n = 20) mean effective power is substantially lower than
standard equal-allocation power calculations suggest.
Effect sizes are reported as standardized mean differences (Cohen's $d$) with
95\% confidence intervals. Given the pilot sample size ($n = 89$), observed
power and required sample sizes for 80\% power are reported alongside null findings.

\subsection{Secondary Analysis: Confidence Change}

Confidence change models mirror the primary belief change sequence and are
estimated using the same OLS/HC3 framework. Two confirmatory and two
exploratory models are specified:

\begin{itemize}
    \item \textbf{conf-M1} (confirmatory): Unadjusted treatment effect on
          confidence change — AI vs.\ human-human, no covariates.
    \item \textbf{conf-M2} (confirmatory): Treatment $+$ reflection quality
          ($z$-scored) — adjusting for engagement depth, mirroring the
          primary M2 specification.
    \item \textbf{conf-M3} (exploratory): Treatment $\times$ reflection
            quality interaction on confidence change — included for
            symmetry with the primary M3 specification; the interaction
            term is not expected to be interpretable given the
            variance-constrained nature of confidence change.
    \item \textbf{belief-ack} (exploratory): Belief change regressed on
        treatment $+$ reflection quality $+$ acknowledgement binary
        (\texttt{acknowledgement\_binary}) — testing whether explicit
        concession behaviour predicts belief change over and above
        reflection quality ($n = 89$, no missingness on this variable).

\end{itemize}

\texttt{belief\_ack} is estimated on the available complete-case sample
($n = 89$ in the current export); results are interpreted as exploratory
and are not included in multiplicity-controlled confirmatory inference.

\subsection{Robustness Checks}

Planned robustness diagnostics for confirmatory models are:
\begin{itemize}
    \item \textbf{Outlier trimming}: Re-estimation after excluding the bottom
          and top 1\% of belief or confidence change values.
    \item \textbf{Leave-one-out (LOO)}: Re-estimation of M2 excluding each
          condition in turn. In runs where the LOO export is empty, no LOO
          numeric claims are carried into inference.
    \item \textbf{VIF diagnostics}: Variance inflation factors computed for
          all multi-predictor models to detect multicollinearity.
    \item \textbf{Shapiro-Wilk}: Residual normality assessed for all models.
\end{itemize}

\subsection{Exploratory Analyses}

Exploratory analyses are not part of primary inference and are reported
as hypothesis-generating. Three categories of exploratory analysis are
conducted.

\subsubsection{Reflection Quality as Mediator}

A qualitative subsample analysis examines the thematic content of
between-round reflections to contextualise the quantitative reflection
quality scores. The subsample is stratified by reflection quality tercile
(low, medium, high) and AI condition, capped at two observations per
debate to avoid within-participant duplication, yielding approximately
$n = 24$ reflections. Reflections are coded for themes including
argument acknowledgement, counter-argumentation style, and evidence
of perspective-taking. This analysis is intended to validate whether
GPT-assigned reflection quality scores correspond to substantively
meaningful differences in engagement, rather than surface-level
textual features.

\subsubsection{Mechanism Analysis: Pushback and Confidence}

As an exploratory module, a mechanism analysis
tests whether critical engagement --- operationalised as pushback
frequency (\texttt{has\_pushback}) --- predicts confidence change.
Pushback is coded from the qualitative subsample (the same
$n \approx 24$ reflections used in the thematic analysis) as the
binary presence of disagreement markers (\textit{but},
\textit{however}, \textit{disagree}, and synonyms). These qualitative
codes are merged with participant-level outcome measures
(\texttt{confidence\_change}, \texttt{belief\_change}) to form the
mechanism dataset.

Pearson correlations between \texttt{has\_pushback} and
\texttt{confidence\_change} are estimated separately within the AI
group and the human-human group. Fisher-$z$ transformation is used
to compute 95\% confidence intervals around each correlation. A
negative correlation within the AI group would suggest that
participants who engaged more critically gained \textit{less}
confidence, contradicting a defensive confidence hypothesis; a
positive correlation would suggest active pushback reinforces
perceived debate success.

Jackknife resampling (leave-one-out) is applied to the AI-group
correlation to assess sensitivity to individual influential
observations. At each iteration, one data point is removed and the
correlation is re-estimated on the remaining subsample, and the
resulting range is reported as a small-sample sensitivity diagnostic
rather than as confirmatory evidence of a stable mechanism effect.
Iterations where pushback or confidence
change variance collapses to zero are excluded.

\subsubsection{Exploring All P1 Variables (Associational)}

An additive exploratory module examines collected player1 (P1) variables that
are outside the pre-specified confirmatory model set. The purpose is
methodological coverage and hypothesis generation, not confirmatory testing.
Where data availability and variance permit, associational checks include
OLS-HC3 models, non-parametric group comparisons, and simple descriptives.
Outputs are reported as exploratory associational diagnostics and are interpreted
jointly with the pushback/jackknife instability checks, without upgrading any
mechanism claim to identified evidence.

Prior to mechanism analysis, qualitative-sample representativeness is
audited by comparing condition proportions in the subsample against
full-population condition proportions. In the current run, the
subsample approximates population condition shares (maximum
$\Delta=0.045$), though modest imbalance remains; mechanism results are
reported as hypothesis-generating and preliminary.


\section{Reporting}

Results are reported with effect sizes and 95\% confidence intervals rather
than $p$-values alone. All models report coefficient estimates,
model-specified standard errors (OLS-HC3 or topic-cluster), and sample sizes.
Exploratory analyses are clearly labelled and
interpreted with appropriate caution. All statistical code, intermediate tables,
and export files are retained to facilitate reproducibility. Cross-chapter
headline numbers are synchronized to Appendix Table~\ref{tab:headline_numbers}.
Chapter 5 follows this reporting hierarchy explicitly: confirmatory outcomes are
reported first, followed by robustness diagnostics, and then exploratory modules
that are interpreted as hypothesis-generating rather than effect-identifying.
