%% Results and Analysis Chapter

\section{Overview}

This chapter reports final results from a descriptive sample of $n=90$ participants across four conditions
(human-human $n=21$, firm $n=28$, balanced $n=20$, open-minded $n=21$), with complete-case model analyses at $n=89$.
The inferential flow is intentionally tiered:
\begin{enumerate}
    \item confirmatory belief-change models,
    \item secondary confidence models,
    \item robustness checks,
    \item exploratory mechanism and qualitative representativeness diagnostics.
\end{enumerate}

All participant-level regression models use robust standard errors, and interpretation
is based on effect size direction and confidence intervals, not on single $p$-values alone.

\pagebreak
\section{At-a-Glance Findings}

\begin{table}[h]
\centering
\caption{Summary of Final Results}
\label{tab:key_findings}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{L{3.8cm}L{7.3cm}L{2.7cm}}
\toprule
Question & Evidence & Decision \\
\midrule
Does AI change belief? & Confirmatory M1, M2, M4, M5 all include 0 in 95\% CI; e.g. M5 $\beta_{AI}=5.233$, CI $[-22.374, 32.840]$, $p=0.710$ & No reliable effect \\
Does AI change confidence? & Secondary confidence models are also non-significant; conf-M2 $\beta_{AI}=5.845$, CI $[-2.565, 14.256]$, $p=0.173$ & No reliable effect \\
Are pooled inferences stable? & Outlier-trim and leave-one-condition-out exports preserve null AI belief effect & Stable null \\
Is pushback mechanism supported? & AI-group pushback-confidence $r=-0.121$, $p=0.707$; jackknife range $[-0.372,-0.003]$ & Weak/uncertain \\
Is qual sample representative? & Qual sample shares deviate from population shares (human-human: $0.500$ vs $0.205$; AI conditions each $0.167$ vs $0.237$--$0.284$) & Not representative \\
\bottomrule
\end{tabular}
\end{table}

\pagebreak
\section{Participant Overview and Data Quality}

\subsection{Condition Composition and Outcome Means}
\label{sec:part10}

Condition-level means for belief change (from \texttt{condition\_overview.csv}) are:
\begin{itemize}
    \item Human-human: $-7.905$
    \item Firm: $1.714$
    \item Balanced: $-0.200$
    \item Open-minded: $1.667$
\end{itemize}

\subsection{Floor/Ceiling Audit}

From \texttt{data\_quality\_floor\_ceiling.csv}:
\begin{itemize}
    \item Belief pre-debate at 0: $20.0\%$
    \item Belief pre-debate at 100: $16.7\%$
    \item Confidence pre-debate at 0: $0.0\%$
\end{itemize}

This supports treating confidence analyses as valid change outcomes in this run
(rather than a degenerate post-only proxy).

\pagebreak
\section{Primary Outcome: Belief Change}

\subsection{Model Family Results}

Table~\ref{tab:primary_models} reports the AI coefficient from the exported primary model table.

\begin{table}[h]
\centering
\caption{Primary Belief Models (AI pooled vs human-human)}
\label{tab:primary_models}
\begin{tabular}{lrrrr}
\toprule
Model & $n$ & $\beta_{AI}$ & 95\% CI & $p$ \\
\midrule
M1 & 89 & 8.695 & $[-9.765, 27.155]$ & 0.3559 \\
M2 & 89 & 6.701 & $[-12.581, 25.984]$ & 0.4958 \\
M3 (exploratory) & 89 & 8.989 & $[-11.393, 29.371]$ & 0.3874 \\
M4 & 89 & 8.334 & $[-10.978, 27.646]$ & 0.3977 \\
M4b (exploratory) & 89 & 6.793 & $[-12.827, 26.413]$ & 0.4974 \\
M4c (exploratory) & 89 & 9.907 & $[-6.698, 26.512]$ & 0.2422 \\
M5 & 89 & 5.233 & $[-22.374, 32.840]$ & 0.7102 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{$R^2$ Comparison Across Primary Models}

Table~\ref{tab:primary_r2_compare} compares variance explained across the primary
model sequence. Reported $\Delta R^2$ values are relative to M1, consistent with
the exported \texttt{primary\_models.csv}.

Model labels used in this subsection:
\begin{itemize}
    \item M1: Treatment only
    \item M2: M1 + Reflection quality
    \item M3: M2 + Interaction (exploratory)
    \item M4: M2 + Stance strength + Topics
    \item M4b: M2 + Arg quality (exploratory)
    \item M4c: M1 + Arg quality without reflection (exploratory)
    \item M5: Full model (M4 + arg quality + rounds + first\_player)
\end{itemize}

\begin{table}[h]
\centering
\caption{$R^2$ Comparison for Primary Model Family}
\label{tab:primary_r2_compare}
\begin{tabular}{lrrr}
\toprule
Model & Type & $R^2$ & $\Delta R^2$ (vs M1) \\
\midrule
M1 & Confirmatory & 0.014 & 0.000 \\
M2 & Confirmatory & 0.049 & 0.034 \\
M3 & Exploratory & 0.071 & 0.057 \\
M4 & Confirmatory & 0.082 & 0.068 \\
M4b & Exploratory & 0.049 & 0.034 \\
M4c & Exploratory & 0.041 & 0.027 \\
M5 & Confirmatory & 0.097 & 0.082 \\
\bottomrule
\end{tabular}
\end{table}

The progression indicates only modest incremental explanatory power across
specifications, with M5 achieving the largest fit increase ($\Delta R^2=0.082$),
while inferential conclusions about the pooled AI coefficient remain unchanged.

\textbf{Primary inference:} no confirmatory model supports a non-zero pooled AI effect on belief change.

\pagebreak
\section{Secondary Outcome: Confidence Change}

\begin{table}[h]
\centering
\caption{Secondary Confidence and Extension Models}
\label{tab:secondary_models}
\begin{tabular}{lrrrr}
\toprule
Model & $n$ & $\beta_{AI}$ & 95\% CI & $p$ \\
\midrule
conf-M1 & 89 & 5.210 & $[-3.021, 13.441]$ & 0.2147 \\
conf-M2 & 89 & 5.845 & $[-2.565, 14.256]$ & 0.1731 \\
conf-M3 (exploratory) & 89 & 6.070 & $[-2.625, 14.764]$ & 0.1712 \\
belief\_ack (exploratory) & 89 & 11.269 & $[-8.980, 31.519]$ & 0.2754 \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Secondary inference:} no statistically reliable pooled AI effect is detected for confidence in this updated run.

\pagebreak
\section{Robustness Checks}

The robustness exports preserve coherence with the main model narrative:
\begin{itemize}
    \item \texttt{robustness\_leave\_one\_condition\_out.csv}: no leave-one-condition variant reverses the primary null AI belief conclusion.
    \item \texttt{robustness\_outlier\_trim.csv}: outlier trimming does not produce a stable non-zero AI belief effect.
    \item Cohort-level descriptors from the summary export remain modest (Cohen's $d=0.286$, condition ANOVA $F=0.488$, $p=0.691$).
\end{itemize}

These checks are included to test model sensitivity under plausible data perturbations,
not to substitute for the confirmatory model family.

\pagebreak
\section{Validation of GPT Quality Scoring Features}

To validate automated quality features used in downstream modeling,
we compare GPT-derived scores against double-rated human judgments
from exported validation samples.

\subsection{Reflection Quality Validation}

Using \texttt{reflection\_validation\_summary.csv} ($n=30$):
\begin{itemize}
    \item Spearman correlation (GPT vs mean human): $\rho=-0.088$, $p=0.643$
    \item Human-only agreement: ICC$(2,1)=0.821$, ICC$(2,k)=0.902$
    \item GPT+human agreement: ICC$(2,1)=0.203$, ICC$(2,k)=0.433$
\end{itemize}

Interpretation: human raters show high internal consistency for reflection quality,
while GPT reflection scores are weakly aligned with the human-ranked ordering in this run.

\subsection{Persuader Argument Quality Validation}

Using \texttt{persuader\_argument\_validation\_summary.csv} ($n=30$):
\begin{itemize}
    \item Spearman correlation (GPT vs mean human): $\rho=0.797$, $p<0.001$
    \item Human-only agreement: ICC$(2,1)=0.715$, ICC$(2,k)=0.834$
    \item GPT+human agreement: ICC$(2,1)=0.725$, ICC$(2,k)=0.888$
\end{itemize}

Interpretation: GPT-derived argument-quality features show strong monotonic alignment
with human ratings and comparable reliability contribution in this sample.

Taken together, the validation evidence is asymmetric by feature family:
argument-quality scoring appears well supported, whereas reflection-quality scoring
should be treated more cautiously for inferential interpretation.

\pagebreak
\section{Exploratory Mechanism and Qualitative Diagnostics}

Using \texttt{mechanism\_validation\_pushback\_confidence.csv}, AI-group mechanism association is:
\begin{align}
r=-0.121, \quad p=0.7074, \quad n=12
\end{align}

The final jackknife range now propagated to the same output file is:
\begin{align}
r_{jackknife} \in [-0.372, -0.003]
\end{align}

This indicates directional negativity but weak inferential certainty at current sample size.

\subsection{Representativeness Caveat}

From \texttt{qual\_sample\_representativeness.csv}, qualitative sample proportions
deviate from full-population condition shares:
\begin{itemize}
    \item Human-human: $0.500$ in qualitative sample vs $0.205$ in population ($\Delta=+0.295$)
    \item Firm: $0.167$ vs $0.284$ ($\Delta=-0.117$)
    \item Balanced: $0.167$ vs $0.274$ ($\Delta=-0.107$)
    \item Open-minded: $0.167$ vs $0.237$ ($\Delta=-0.070$)
\end{itemize}

Hence exploratory qualitative/mechanism findings remain hypothesis-generating.

\pagebreak
\section{Synthesis}

The updated evidence chain is internally coherent:
\begin{enumerate}
    \item Confirmatory belief models: null pooled AI effect.
    \item Secondary confidence models: null pooled AI effect in this run.
    \item Robustness checks: no sensitivity configuration overturns the confirmatory null.
    \item Exploratory mechanism/qualitative outputs: informative but non-adjudicative due to uncertainty and representativeness limits.
\end{enumerate}

Overall, the current dataset supports a conservative conclusion: under these model specifications
and this sample, pooled AI opponent type does not produce a reliable shift in either belief change
or confidence change, while exploratory diagnostics remain useful for next-step hypothesis design.
For cross-chapter consistency, headline numbers in this chapter are anchored to
Appendix Table~\ref{tab:headline_numbers}.
