%% System Design & Architecture Chapter
%% This file is included in main.tex as a chapter

\section{Introduction}

This chapter documents only the system decisions that directly support internal
validity and analyzable persuasion data. Implementation details not required for
this argument are moved to Appendix A.

\section{Information Hiding as an Experimental Control}

The core design principle is information hiding: during active debate rounds,
participants are not told whether the opponent is human or AI. Attribution is
measured only after the debate ends. This reduces source-label effects during
argument evaluation and supports cleaner interpretation of observed belief and
confidence trajectories.

\section{Turn State and Between-Round Data Integrity}

Debates are enforced as a turn-gated sequence: argument submission, then
between-round measurement, then progression to the next turn. The server tracks
turn state explicitly and rejects out-of-order transitions. This protocol-level
state machine is the main data-integrity mechanism for the thesis because it
anchors each measurement to a specific interaction boundary and reduces temporal
ambiguity in round-level analysis.

The between-round block captures three key process measures at each boundary:
belief update, confidence update, and short reflection/acknowledgement text.
Because collection is embedded in turn progression, missingness and post-hoc
recall error are reduced relative to end-only survey capture.
Each argument turn is also constrained by a fixed 500-character input limit,
which standardizes message length across conditions and supports fairer
cross-condition comparison of per-round submissions.

\section{Research-Relevant Schema Fields}

MongoDB stores debate records with the fields needed for the analysis pipeline.
For this thesis, the most important schema groups are shown in
Table~\ref{tab:research_schema_core}.

\begin{table}[htbp]
\centering
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{p{3.0cm}p{5.3cm}p{6.1cm}}
\toprule
\textbf{Entity} & \textbf{Core Fields} & \textbf{Research Purpose} \\
\midrule
Debate metadata & {\ttfamily topicId, condition, status, currentRound} & Define condition assignment and round boundaries for modeling \\
Arguments & {\ttfamily arguments[stance, text, round, submittedBy, timestamp]} & Preserve ordered claim/counterclaim content \\
Belief trajectory & {\ttfamily beliefHistory[belief\allowbreak Value, confidence, influence, round, userId]} & Estimate round-level belief and confidence dynamics \\
Reflection & {\ttfamily reflections[paraphrase, acknowledgement, round, userId, timestamp]} & Capture engagement and concession signals \\
Post-debate survey & {\ttfamily player1Opponent\allowbreak Perception, confidence, cues} & Evaluate attribution and endpoint self-reports \\
\bottomrule
\end{tabular}
\caption{Core schema fields used in this thesis}
\label{tab:research_schema_core}
\end{table}

\section{Appendix Boundary}

WebSocket event catalogs, transport lifecycle details, authentication/session
engineering, operational scalability notes, and extended error-handling flows are
moved to Appendix A to keep this chapter focused on methodological validity.

\section{Methodological Traceability from Platform to Models}

To make Chapter 3 analytically complete without reproducing implementation code,
this section maps each core platform mechanism to the corresponding modeling
consequence used in later chapters.

\begin{itemize}
    \item \textbf{Hidden opponent identity during rounds} $\rightarrow$ supports cleaner estimation of
	    AI-vs-human contrasts by reducing immediate source-label demand effects during argument evaluation.
    \item \textbf{Turn-gated state transitions} $\rightarrow$ preserves round order validity and makes
	    per-round trajectories interpretable as temporally aligned process data.
    \item \textbf{Mandatory between-round measurement} $\rightarrow$ yields structured belief/confidence
	    trajectories and reflection/acknowledgement signals for mechanism analysis.
    \item \textbf{Fixed turn-length constraint (500 characters)} $\rightarrow$ reduces variance induced by
	    extreme response-length differences and keeps interaction opportunities more comparable across conditions.
    \item \textbf{Schema fields shown in Table~\ref{tab:research_schema_core}} $\rightarrow$ provide a
	    direct bridge from platform events to the confirmatory model variables reported in Chapters 4 and 5.
\end{itemize}

In short, the architecture is not presented as generic software documentation;
it is presented as a validity argument. The implementation choices above define
what causal and associational claims are supportable from the collected data,
while Appendix A retains the engineering details needed for reproducibility.
The implementation repository and companion technical materials are archived at
the project source repository \cite{debateplatform_repo}.

\section{Conclusion}

System design contributes to the study primarily through three mechanisms:
information hiding, enforced turn-state integrity, and structured between-round
measurement capture. Together these make the resulting data suitable for the
belief/confidence model sequence and trajectory-based interpretation used in
later chapters.
